---
title: "Foresite Emulation: an example."
author: Luis Damiano
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The following code provides an illustration on how to use the package:

* Draw LHSs from an empirical dataset of soil measurements.
    * The function `lhs_soil` expands soil measurements at each map unit,
      identified by `unitKey`, into a soil profile with cuts at depth
      `soilLayers`.
* Draw LHSs from an empirical dataset of weather measurements.
    * The function `lhs_weather` takes a dataset in long format.
* Draw LHSs from an empirical dataset of management decisions.
    * If such an empirical dataset does not exist, you can use
      `management_population` to generate a large sample of decisions mimicking
      all possible distributions.
* Combine all draws into runs.
    * The elements of the largest soil/weather/management samples are run only once.
    * The elements of the other two samples are recycled so that each is used
      in approximately the same number of runs.

The example is not run in the vignette, do it by hand to see the results.

```{r setup, eval = FALSE}
library(ForesiteEmu)

# Use data.table under the hook for performance as the mgmt dataset is large.
fread       <- data.table::fread
fwrite      <- data.table::fwrite

# Settings
seed        <- 9000
nSoilLHSs   <- 200
nWeatherLHS <- 30
nMgmtLHSs   <- 50000

# Soil sampling -----------------------------------------------------------
# Prepare empirical data
ssurgoPath  <- system.file("extdata", "ssurgo_ia169.txt", package = "ForesiteEmu")
ssurgoData  <- read.table(ssurgoPath, sep = "\t", header = TRUE)
ssurgoCols  <- c(
  "cokey", "hzdept_r", "hzdepb_r", "slope", "kffact", "dbthirdbar_r", 
  "ph1to1h2o_r", "wthirdbar_r", "wfifteenbar_r", "awc_r", "claytotal_r", 
  "sandtotal_r", "ksat_r", "om_r", "ll_r"
)

x           <- na.omit(ssurgoData[, ssurgoCols])

# Sampling settings
unitKey     <- "cokey" # Variable identifying 
soilLayers  <- c(20, 40, 60, 80, 100, 150, 200)
soilLHSs    <- lhs_soil(x, unitKey, soilLayers, nSoilLHSs, seed)

fwrite(x = soilLHSs, file = "./soil_sample.csv")

# Weather sampling --------------------------------------------------------
# Prepare empirical data
weatherPath <- system.file(
  "extdata", "weather_ia169_1980_2018.txt", package = "ForesiteEmu"
)
weatherData <- read.table(file = weatherPath, sep = "\t", header = TRUE)[-1]
weatherCols <-
  c("year", "yday", "dayl", "prcp", "srad", "swe", "tmax", "tmin", "vp")

colnames(weatherData) <- weatherCols
x           <- na.omit(weatherData)

# Sampling settings
weatherLHSs <- lhs_weather(x, nWeatherLHS, seed)
fwrite(x = weatherLHSs, file = "./weather_sample.csv")

# Management sampling -----------------------------------------------------
# Simulate empirical data (we save it to a file for future use)
mgmtPath    <- "./mgmt_population.csv"

if (!file.exists(mgmtPath))
  fwrite(management_population(m = 2), mgmtPath)

x           <- fread(mgmtPath)

# Sampling settings
mgmtLHSs    <- lhs_management(x, nMgmtLHSs, seed)

fwrite(x = mgmtLHSs, file = "./mgmt_sample.csv")

# Generate runs -----------------------------------------------------------
set.seed(seed)

nRuns       <- max(nSoilLHSs, nWeatherLHS, nMgmtLHSs)
runs        <- data.frame(
  run_id               = 1:nRuns,
  soil_sample_id       = sample(rep(1:nSoilLHSs, len = nRuns)),
  weather_sample_id    = sample(rep(1:nWeatherLHS, len = nRuns)),
  management_sample_id = sample(rep(1:nMgmtLHSs, len = nRuns))
)

fwrite(x = runs, file = "./run_samples.csv")
```
